---
title: "Analysis of 'but not both' experiment"
author: "jdegen"
date: "Dec 14, 2017"
output: html_document
---

## Goal

Figure out, for each case, the best "but not both" sentence to run in the scalar implicature experiment (ie, analogous to the "some" paper, comparing original sentence to one with potentially implicated content locally explicit). Even though most of the sentences sound terrible with "but not both" -- which already tells us that the "and" version usually is NOT a salient alternative.

## What we initially wanted to determine
What is the best "but not both" sentence selection criterion, of the three we have? Neele's initial pilot analysis suggested that the modal response was best for 25 of 30 sentences. In general: the modal response seems best for cases that receive only few different responses. For cases with more diversity in responses, selecting the case with the greatest proportion of "good location" responses seems to work best.  

Questions that need to be answered:

1. Is it even worth running sentences with too low of a proportion of "good location" responses? What is a reasonable cutoff point for proportion of "good location" responses?
2. What is the cutoff point for number of different lcoations of "but not both" above which we should deviate from the modal response selection method?
3. Which cases need to be looked at manually, if any?

Selection criteria:

1. **Include** cases where all three response selection methods yield the same result and **select modal response**.
2. **Exclude** cases with lower than XXX (10-20%)% "good location" responses. (If 1. and 2. are in conflict, include or exclude? Hand-check some responses to come up with general rule)
3. **Include** cases with higher than XXX (70-80)% "good location" responses. If selection methods are in conflict:
  - if fewer than (3(oder2))XXX different responses were generated, **select modal response**
  - if more than (3)XXX different responses were generated, **exclude**
  - else, **select proportion response**

## Columns

- **tgrep.id/TGrepID:** unique TGrep2 ID
- **entire.sentence/EntireSentence:** original sentence, without 'but not both' insertion
- **Num_different_BNBForms/NumBNBForms**: number of different locations for "but not both" that were generated
- **Proportion_goodLoc_perBNB/PropGoodLocation:** proportion of participants that rated the case as having a good location for "but not both" -- Neele, what is "the case" here? Ie if modal response etc differ, what does this column code?
- **modal_response/ResponseModal:** the most frequently given response
- **ResponseFreqGoodLocation/ResponseFreqGoodLocation:** the response with greatest frequency of "good location" responses
- **ResponsePropGoodLocation/ResponsePropGoodLocation:** the response with greatest proportion of "good location" responses
- **SentenceOverlap/SentenceOverlap:** do all three methods for selecting the best response yield the same result?
- **MoreThanOneModalResponse/MoreThanOneModalResponse**: are there more than one BNB Sentences that are most frequent?
-**MoreThanOneFreqResponse/MoreThanOneFreqResponse**: are there more than one BNB Sentences with most frequency of "good location" responses?
-**MoreThanOnePropResponse/MoreThanOnePropResponse**: are there more than one BNB Sentences with greates proportion of "good location" responses?
-**ResponseBNBClosestToOrModal/ResponseBNBClosestToOrModal**: if MoreThanOneModalResponse is True, this encodes the most frequent BNB Sentence where or is closer to but not both, but not directly followed by it.
-**ResponseBNBClosestToOrFreq/ResponseBNBClosestToOrFreq**: if MoreThanOneFreqResponse is True, this encodes the BNB Sentence with most frequent "good location" responses where or is closer to but not both, but not directly followed by it.
**-ResponseBNBClosestToOrProp/ResponseBNBClosestToOrProp**: if MoreThanOnePropResponse is True, this encodes the BNB Sentence with highest proportion of "good location" responses where or is closer to but not both, but not directly followed by it.
**-BestResponse**: This will encode the best BNB Sentence, extracted by evaluatin different criteria

Load packages, read the data, rename columns, compute sentence length (will be twice as long for sentences with two best responses), create an empty new column with final sentence to use.
```{r}
library(tidyverse)
library(magrittr)
library(knitr)
d = read.table("../data/final.csv",header=T,sep="\t",quote="")
colnames(d) = c("TGrepID","EntireSentence","NumBNBForms","AllBnBs","PropGoodLocation","ResponseModal","ResponseFreqGoodLocation","ResponsePropGoodLocation","SentenceOverlapMFP", "MoreThanOneModalResponse","MoreThanOnePropResponse","MoreThanOneFreqResponse", "MoreThanOneResponse","ResponseBNBClosestToOrModal", "ResponseBNBClosestToOrFreq", "ResponseBNBClosestToOrProp", "BNBClosestToOr")
d %<>%
       mutate_at(c("ResponseModal","ResponseFreqGoodLocation","ResponsePropGoodLocation","BNBClosestToOr"),funs(as.character(.)))
d %<>%
       mutate_at(c("ResponseModal","ResponseFreqGoodLocation","ResponsePropGoodLocation","BNBClosestToOr"),funs(trimws(.)))
d$SentenceLength = nchar(as.character(d$EntireSentence))
d[is.na(d$BNBClosestToOr),]$BNBClosestToOr = d[is.na(d$BNBClosestToOr),]$ResponseModal
d$CompleteSentenceOverlap = ifelse(d$ResponseModal == d$ResponseFreqGoodLocation, ifelse(d$ResponseModal == d$ResponsePropGoodLocation, ifelse(d$ResponseModal == d$BNBClosestToOr,T,F),F),F)

d$BestResponse = ""
```
There are `r {nrow(d)}` cases.

## Selection process

### **Include** cases where all four response selection methods yield the same result and **select modal response**.

In `r {nrow(d[d$CompleteSentenceOverlap == T,])}` of the cases the same "but not both" sentence is chosen regardless of choice criterion. Some of those sentences:

```{r}
table(d$CompleteSentenceOverlap,d$NumBNBForms)
d[d$CompleteSentenceOverlap == T,]$BestResponse = d[d$CompleteSentenceOverlap == T,]$ResponseModal
sample(d[d$CompleteSentenceOverlap == T,]$ResponseModal,5)
```

Compute number of cases with overlap in any combination of 3 responses:

```{r}
d$SentenceOverlapMFP  = ifelse(d$ResponseModal == d$ResponseFreqGoodLocation, ifelse(d$ResponseModal == d$ResponsePropGoodLocation, ifelse(d$ResponseModal != d$BNBClosestToOr,T,F),F),F)
d$SentenceOverlapMFC  = ifelse(d$ResponseModal == d$ResponseFreqGoodLocation, ifelse(d$ResponseModal == d$BNBClosestToOr, ifelse(d$ResponseModal != d$ResponsePropGoodLocation,T,F),F),F)
d$SentenceOverlapMPC = ifelse(d$ResponseModal != d$ResponseFreqGoodLocation, ifelse(d$ResponseModal == d$BNBClosestToOr, ifelse(d$ResponseModal == d$ResponsePropGoodLocation,T,F),F),F)
d$SentenceOverlapFPC = ifelse(d$ResponsePropGoodLocation == d$ResponseFreqGoodLocation, ifelse(d$ResponsePropGoodLocation == d$BNBClosestToOr, ifelse(d$ResponseModal != d$ResponsePropGoodLocation,T,F),F),F)
```

Cases with MFP overlap -- choose closest to "or":
```{r}
table(d$SentenceOverlapMFP,d$NumBNBForms)
kable(sample(d[d$SentenceOverlapMFP,c("TGrepID","NumBNBForms","ResponseModal","BNBClosestToOr")]))
```

Cases with MFC overlap -- choose closest to "or":
```{r}
table(d$SentenceOverlapMFC,d$NumBNBForms)
kable(sample(d[d$SentenceOverlapMFC,c("TGrepID","NumBNBForms","ResponseModal","ResponsePropGoodLocation")]))
```

Cases with MPC overlap -- choose closest to "or":
```{r}
table(d$SentenceOverlapMPC,d$NumBNBForms)
kable(sample(d[d$SentenceOverlapMPC,c("TGrepID","NumBNBForms","ResponseModal","ResponseFreqGoodLocation")]))
```

Cases with FPC overlap -- choose closest to "or":
```{r}
table(d$SentenceOverlapFPC,d$NumBNBForms)
kable(sample(d[d$SentenceOverlapFPC,c("TGrepID","NumBNBForms","ResponseModal","BNBClosestToOr")]))
```

Compute number of cases with overlap in modal response and closest to "or":

```{r}
d$SentenceOverlapMC  = ifelse(d$ResponseModal  == d$BNBClosestToOr,T,F)
table(d$SentenceOverlapMC)
kable(sample(d[d$SentenceOverlapMPC,c("TGrepID","NumBNBForms","ResponseModal","ResponseFreqGoodLocation")]))

```

Show random sample of cases with no overlap in modal response and closest to "or" (conditioning on num of BNB forms being some number):
```{r}
kable(arrange(sample(d[d$ResponseModal != d$BNBClosestToOr & d$NumBNBForms == 2,c("TGrepID","SentenceLength","NumBNBForms","ResponseModal","BNBClosestToOr")]),SentenceLength))

```

Selection:
When there is overlap in at least modal and closest to "or", choose that best response. When there is no overlap, we do pairwise sentence norming (see below).
```{r}
d[d$SentenceOverlapMC,]$BestResponse = as.character(d[d$SentenceOverlapMC,]$ResponseModal)
```

In how many cases of no overlap between modal and closest to "or" does modal have more than one response? Are the closest to "or" responses reasonable in all those cases?

```{r}
table(d[!d$SentenceOverlapMC,]$MoreThanOneModalResponse)
kable(d[!d$SentenceOverlapMC & d$MoreThanOneModalResponse,c("BNBClosestToOr","ResponsePropGoodLocation")])

kable(d[!d$SentenceOverlapMC & d$MoreThanOneModalResponse & d$MoreThanOneFreqResponse & d$MoreThanOnePropResponse ,c("BNBClosestToOr","ResponsePropGoodLocation")])
```

In Exp. 2 we norm comparisons between sentences. For the cases that don't have more than one modal response, we compare the modal response to the closest to "or" (564). For the cases that do have more than one but don't have more than one freq response, we compare the freq response to the closest to "or" (29). For the cases that do have more than one but don't have more than one prop response, we compare the prop response to the closest to "or" (8). For the remaining 16 we hand-selected the best response.
```{r}
table(d[!d$SentenceOverlapMC,]$MoreThanOneModalResponse)
table(d[!d$SentenceOverlapMC & d$MoreThanOneModalResponse,]$MoreThanOneFreqResponse)
table(d[!d$SentenceOverlapMC & d$MoreThanOneModalResponse & d$MoreThanOneFreqResponse,]$MoreThanOnePropResponse)

tonorm = d %>%
  filter(!SentenceOverlapMC & (!MoreThanOneModalResponse | !MoreThanOneFreqResponse | !MoreThanOnePropResponse))
tonorm$ComparisonSentence = ""
tonorm[!tonorm$MoreThanOneModalResponse,]$ComparisonSentence = as.character(tonorm[!tonorm$MoreThanOneModalResponse,]$ResponseModal)
tonorm[tonorm$MoreThanOneModalResponse & !tonorm$MoreThanOneFreqResponse,]$ComparisonSentence = as.character(tonorm[tonorm$MoreThanOneModalResponse & !tonorm$MoreThanOneFreqResponse,]$ResponseFreqGoodLocation)
tonorm[tonorm$MoreThanOneModalResponse & tonorm$MoreThanOneFreqResponse & !tonorm$MoreThanOnePropResponse,]$ComparisonSentence = as.character(tonorm[tonorm$MoreThanOneModalResponse & tonorm$MoreThanOneFreqResponse & !tonorm$MoreThanOnePropResponse,]$ResponsePropGoodLocation)
write.table(tonorm %>% select(TGrepID,EntireSentence,BNBClosestToOr,ComparisonSentence), file="../data/tonorm_round2.csv",row.names=F,quote=F,sep="\t")
```

## Exp 2 norming results

Load the norming results, assign best responses.

```{r}
norms = read.csv("../../../2_butnotboth_norm/results/data/bestresponses.csv")
colnames(norms) = c("TGrepID","Response","N")
tonorm_withresponses = left_join(tonorm,norms,by=c("TGrepID"))
tonorm_withresponses$BestResponse = ""
tonorm_withresponses[tonorm_withresponses$Response == "BNBClosestToOr",]$BestResponse = tonorm_withresponses[tonorm_withresponses$Response == "BNBClosestToOr",]$BNBClosestToOr
tonorm_withresponses[tonorm_withresponses$Response == "ComparisonSentence",]$BestResponse = tonorm_withresponses[tonorm_withresponses$Response == "ComparisonSentence",]$ComparisonSentence
row.names(d) = d$TGrepID
d[tonorm_withresponses$TGrepID,]$BestResponse = tonorm_withresponses$BestResponse
write.table(d[,c("TGrepID","EntireSentence","BestResponse")],"../data/bestresponses.csv",row.names=F,quote=F,sep="\t")
write.table(d[d$BestResponse=="",c("TGrepID","EntireSentence","BestResponse")],"../data/bestresponse_empty.csv",row.names=F,quote=F,sep="\t")
```

Get results for paper.

```{r}
table(tonorm_withresponses[!tonorm_withresponses$SentenceOverlapMC,c("MoreThanOneModalResponse","Response")])
prop.table(table(tonorm_withresponses[!tonorm_withresponses$SentenceOverlapMC,c("MoreThanOneModalResponse","Response")]),mar=c(1))

table(tonorm_withresponses[!tonorm_withresponses$SentenceOverlapMC & tonorm_withresponses$MoreThanOneModalResponse,c("MoreThanOneFreqResponse","Response")])
prop.table(table(tonorm_withresponses[!tonorm_withresponses$SentenceOverlapMC & tonorm_withresponses$MoreThanOneModalResponse,c("MoreThanOneFreqResponse","Response")]),mar=c(1))

table(tonorm_withresponses[!tonorm_withresponses$SentenceOverlapMC & tonorm_withresponses$MoreThanOneModalResponse & tonorm_withresponses$MoreThanOneFreqResponse,c("MoreThanOnePropResponse","Response")])
prop.table(table(tonorm_withresponses[!tonorm_withresponses$SentenceOverlapMC & tonorm_withresponses$MoreThanOneModalResponse & tonorm_withresponses$MoreThanOneFreqResponse,c("MoreThanOnePropResponse","Response")]),mar=c(1))

```



### DEPRECATED -- **Exclude** cases with lower than XXX (10-20%)% "good location" responses? Nope.

The following has been made obsolete by the decision to run a second norming round.

For the cases that don't include the already included cases from 1, what is the distribution of proportion of good locations?

```{r}
dsub = d %>%
  filter(!d$SentenceOverlap)
ggplot(dsub, aes(x=PropGoodLocation)) +
  geom_histogram()
```

Is there a correlation between proportion of good location and number of BNB forms? Yes, a weak one.

```{r}
ggplot(dsub, aes(x=NumBNBForms,y=PropGoodLocation)) +
  geom_jitter() +
  geom_smooth(method="loess")
cor(dsub$NumBNBForms,dsub$PropGoodLocation)
```

Is there a correlation between sentence length and both a) proportion of good location and b) number of BNB forms? Yes, a weak one.

```{r}
ggplot(dsub, aes(x=SentenceLength,y=PropGoodLocation)) +
  geom_point() +
  geom_smooth(method="loess")

ggplot(dsub, aes(x=SentenceLength,y=NumBNBForms)) +
  geom_jitter() +
  geom_smooth(method="loess")
```

Start by looking at the cases with large number of BNB forms -- exlude?

```{r}
nrow(dsub[dsub$NumBNBForms > 4,])
kable(dsub[dsub$NumBNBForms > 4,c("ResponseModal","ResponseFreqGoodLocation","ResponsePropGoodLocation")])
kable(arrange(d[c("NumBNBForms","ResponseModal","ResponseFreqGoodLocation","ResponsePropGoodLocation","BNBClosestToOr")],NumBNBForms))
```


### NExt steps

Before using above selection criteria, Neele will generate new "best response" column that is result of choosing the case where "but not both" is closest to "or" but doesn't come right after it. Then we do the above comparison again.